---
title: "STA 220 - Data and Web Technologies for Data Analysis - Lab 3"
params:
   term: "Winter 2021"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```


Consider  the following set of probabilities 0.5, 0.6, 0.7, 0.8 and 0.9 for the example in Lecture 3.

>- What should we expect to observe?

>- What happen if we consider 100 heads in a row?


## Hint


Use the functions of Lecture 3:



```{r }
simtosses <- function(nsim, ntoss, probH = 1/2) {
    matrix(sample(0:1, ntoss * nsim, replace = TRUE,
                  prob = c(1 - probH, probH)),
           ncol = nsim)
}
```

```{r fmaxrl}
fmaxrl <- function(x, outcome = NULL) {
  xr <- rle(x)
  if (!is.null(outcome)) {
    is_outcome <- (xr$values == outcome)
    xr$lengths <- xr$lengths[is_outcome]
  }
  if (length(xr$lengths) == 0L) 0 else max(xr$lengths)
}
```

```{r }
nsim <- 10000
ntoss <- 100
pH <- c(0.5, 0.6, 0.7, 0.8, 0.9)
res <- matrix(nrow = nsim, ncol = length(pH))
colnames(res) <- pH
for (i in seq_along(pH)) {
    res[, i] <- apply(simtosses(nsim, ntoss, probH = pH[i]), 2, fmaxrl)
}
muhat <- apply(res, 2, mean)
sigmahat <- apply(res, 2, sd)
se.muhat <- sigmahat / sqrt(nsim)
round(rbind(muhat, se.muhat), 3)
```




# We can use the function apply

```{r }
apply(res, 2, quantile, probs = c(0, 0.05, 0.25, 0.5, 0.75, 0.95, 1))
```


```{r biasedcoin.plot, eval = FALSE}
opar <- par(mar = c(5.1, 4.1, 1.1, 1.1))
boxplot(res, horizontal = TRUE,
        xlab = "Maximum Run Length", ylab = "Probability of Heads")
par(opar)
```
---

```{r, ref.label="biasedcoin.plot", echo=FALSE, fig.align = "center", fig.width = 10, fig.height = 5, fig.cap = paste0("Maximum run lengths in ", format(nsim, big.mark = ","), " simulations of ", ntoss, " tosses of a biased coin.")}
```

# If we consider 100 heads in a row

With the probability of heads set to 0.9, the probability of getting 100 heads in 100 (independent) tosses of the (biased) coin is pretty small.

```{r}
0.9^100
```

However, if we repeat the experiment 10,000 times, the probability that we get 100 heads at least once is not so small.

```{r}
1 - (1 - 0.9^100)^10000
```