---
title: "STA 220 - Data and Web Technologies for Data Analysis - Lab 7"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Load library packages

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(moderndive)# Tidyverse-Friendly Introductory Linear Regression
library(broom)# Convert Statistical Objects into Tidy Tibbles 
```

## Preview of a model

`starwars` is a tibble with 87 rows and 14 varibales. See `?dplyr::starwars`

```{r}
mymodel <- lm(mass ~ height, data = starwars)
mymodel
print("XMXMXMXMXMXMXMXMXM")
summary(mymodel)
```


## Data

data are from the `moderndive` package.  [_Modern dive_](https://moderndive.com/) by Ismay and Kim.  

```{r}
evals_ch5 <- evals %>% 
  select(ID, score, bty_avg, age)

evals

evals_ch5
```

# Summary

```{r}
evals_ch5 %>% 
  summary()
```

```{r}
skimr::skim(evals_ch5)# Skim a data frame, getting useful summary statistics
```
## Correlation

`moderndive::get_correlation`: Get correlation value in a tidy way.

```{r}
evals_ch5 %>% 
  get_correlation(score ~ bty_avg)
```

Get correlation of mass as a prediction of height

```{r}
starwars %>% 
  filter(mass < 500) %>% # also filter NA
  summarise(cor = cor(height, mass))
```

### visualize correlation

```{r}
evals_ch5 %>% 
  ggplot(aes(score, age)) +
  geom_jitter() +
  geom_smooth(method = lm)
```

```{r}
evals_ch5 %>% 
  get_correlation(age ~ score)
```

## summary model 

> For every increase of 1 unit increase in bty_avg, there is an associated increase of, on average, 0.067 units of score. from [_ModenDive_](https://moderndive.com/5-regression.html)

```{r}
# Fit regression model:
score_model <- lm(score ~ bty_avg, data = evals_ch5)
```


## Regression Table


```{r}
tidy(score_model)# broom: Tidy summarizes information about the components of a model. 
get_regression_table(score_model)# moderndive: Output regression table for an lm() regression in "tidy" format. 

```

### Regression Summary

```{r}
glance(score_model)# broom: Glance accepts a model object and returns a tibble::tibble() with exactly one row of model summaries. 
get_regression_summaries(score_model)# moderndive: Output scalar summary statistics for an lm() regression in "tidy" format.
```


### Regression points (and residuals)

```{r}
augment(score_model)# broom: Augment accepts a model object and a dataset and adds information about each observation in the dataset.
get_regression_points(score_model)# moderndive: Output information on each point/observation used in an lm() regression in "tidy" format.
```


